Parsing Tweets into Universal Dependencies

### 2017-07-11 paper.tex
linguistic phenomena 1

@@

\section{Data Annotation}
% Linguistic phenomena
	% word level
		% RT
		% Hashtag
		% Mention, urls
		% Emoticons
		% Partially Elided tokens
	% twitter level
		% RT, Answer
		% Parataxis	
% POS
	% Acronym priority	
% Depenency
	% Concentrate on NON-UD phenomena
% Statistics
	% Scale
	% Inter-annotator agreement
	% Proportion
		% how many hashtags, mentions, ....
		% how many acornyms
\subsection{}

@@

\subsection{POS Tagging}
Different from standard sentences, special constructions in tweets exist in both the whole tweet-level and token-level.

\subsubsection{Token-level Unconventional Part-of-Speech}
In addition to standard language, tweets have the following specific constructions:
\begin{itemize}
\item Discourse marker, URL, hashtag, emoticon
	\begin{itemize}
		\item RT, http://bit.ly/xyz, \#acl, :-), @user
	\end{itemize}
\item Acronym
	\begin{itemize}
		\item wtf (what the fuck), smh (shake my hand), mfw (my face when), ima (i am going to), rn (right now), af (as fuck)
	\end{itemize}
\item Contraction
	\begin{itemize}
		\item Contraction with apostrophe: There’s, he’s, book’ll, buy’em
		\item Contraction without apostrophe: gonna, tryna, gimme, Iv, im, aint, whatis
	\end{itemize}
\item Partially Elided Tokens
\end{itemize}

Gimpel et al. \shortcite{Gimpel:2011:PTT:2002736.2002747} proposed a set of part-of-speech tags that handles most of these cases. 
We are inspired by their work but argue that, first different abbreviation can have different syntactic functions. 
Like “mfw” is usually followed by an adverbial clause and “ima” is usually followed by a clausal complement. 
It is not reasonable to treat them in the same part-of-speech. 
In this paper, when annotating the POS tagging for abbreviations, we first try to recover their original forms, then use the POS of the core-word as the POS for the abbreviation.
Second, four special POS tags (S, L, M, Y) were designed to handle contraction words in Gimpel et al. \shortcite{Gimpel:2011:PTT:2002736.2002747}. Major concern of designing such tags is to minimize the effort of tokenization. 
However, contractions of common nouns and pronouns are casted into the same category which increase the difficulty of distinguishing their syntactic function (say, there's and book'll are treated with the same syntactic function). What's more, only a small proportion of words can be categorized into these tags (2.7 \% in total), which cast a doubt of the usefulness of these certain tags. In this paper, we believe such contraction can be properly handled by tokenization module, so we suggest to tokenize the contraction word and annotate POS tag accordingly.
Besides the contraction that be conventionally tokenized, tweets also witness a set of unconventional contraction like iv (I've), whatis (what is). In this paper, we follow the same idea of annotation abbreviation to handle the unconventional contractions and use the POS of core word of the original form as their POS.
Third, special POS was designed to handle emoticon in Gimpel et al. \shortcite{Gimpel:2011:PTT:2002736.2002747}. However, in most cases, emoticon plays the same role as most of the symbolic tokens. In this paper, we follow the UD guideline to annotate the emoticon as symbol (SYM).
At last, it’s arguable that some of the hashtags, URLs can work as a nominal in tweets. Whether treating them as the same part-of-speech or different ones according to their context is an open question. A preliminary survey on the standard UD English data shows that URL, email address are all tagged as the foreign language (X), so we also tag them as X and leave the disambiguation of their syntactic function to the annotation of parse tree.

>>>>

%%%%%%%
\usepackage{graphicx}
\newcommand{\heart}{\ensuremath\heartsuit}

@@

\section{Data}
% Linguistic phenomena
	% word level
		% RT
		% Hashtag
		% Mention, urls
		% Symbols and Emoticons
		% Truncated tokens
	% twitter level
		% RT, Answer
		% Parataxis	
% POS
	% Acronym priority	
% Depenency
	% Concentrate on NON-UD phenomena
% Statistics
	% Scale
	% Inter-annotator agreement
	% Proportion
		% how many hashtags, mentions, ....
		% how many acornyms

\subsection{Linguistic Phenomena of Twitter}
As the representative of the web language, different from the standard languages such as newswire, twitter has its own linguistic phenomena, that could be categorized into word level and structure level phenomena, and we will discuss them separately.

\subsubsection{Token Level}

There are special tokens that will appear only or much more frequently in tweets, including

\begin{itemize}
\item Retweet discourse marker: RT
\item URL: http://bit.ly/xyz
\item Hashtag: \#ACL
\item At-mentioned username: @user
\item Emoticon, emoji and symbol: :), - - - - $\langle$$\langle$$\langle$, :-), …
\item Acronym: wtf (what the fuck), smh (shake my hand), mfw (my face when), ima (i am going to), rn (right now), af (as fuck)
\item Contraction: he's, buy'em, gonna, trma, gimme, im
\item Truncated token: because each tweet can have only 140 characters, excessive characters will be cut off, and usually there could be token in the last of the tweet that is partially elided
\end{itemize}

Consider the following example

\begin{center}
RT @Yijia : \#ACL2017 im heading to Canada for ACL can u gimme some money :) http://url1 I would love to do it with th …
\end{center}

From the example, we argue that most of the the retweet discourse markers, URLs, hashtags, usernames and the emoticons/emojis/symbols usually do not have clear syntactic functions in the tweet, and therefore we should not include them in the analysis, and treat them all equally as non-syntactic tokens. Especially, for truncated tokens, although it might be possible to recover the original tokens from the context in some cases, and we can infer that the truncated token ``th" could be probably the token ``the" it is extremely hard to further predict the rest of the elided sentence. For the simplicity of the annotation, we will consider all the truncated tokens as non-syntactic tokens.

For acronyms and contractions, it is obviously better if we could recover their original forms before analyzing the whole tweet, and we believe that all of these tokens will always be syntactically part of the tweet.

However, consider another example

\begin{center}
@Yi u gonna $\heart$ it . \#NAACL will be \#Awesome . Check my new paper with @nlpnoah on http://url2 RT it !!
\end{center}

Similar in the analysis of Kong et al. \shortcite{kong-EtAl:2014:EMNLP2014}, in some cases, retweet marker (RT), URLS (http://url2), hashtags (\#NAACL, \#Awesome), usernames (@Yi, @nlpnoah) and emoticons ($\heart$) can also have syntactic functions, and we need to take them into the account in the syntactic analysis.

We define acronyms and contractions as full syntactic tokens, truncated tokens as zero syntactic tokens, and all of the other tokens as partial syntactic tokens.

\subsubsection{Structure Level}
Besides the token level phenomena of twitter, there are also structural patterns appearing in the twitter.

\begin{itemize}
\item Retweet structure: RT @user :  $\langle$ tweet content $\rangle$ is a typical structural pattern when a user is retweeting from other users
\item Parataxis: Very often, one tweet is comprised of many sentences or phrases without any delimiting punctuations, such as $\langle$ sentence1 $\rangle$ $\langle$ phrase1 $\rangle$ $\langle$ sentence2 $\rangle$ ...
\end{itemize}

We treat ``RT @user" as the structural patterns in tweet and keep their annotations consistent across tweets. For parataxis case

%vocative X pos tags??

\subsection{POS Tagging}
Gimpel et al. \shortcite{Gimpel:2011:PTT:2002736.2002747} proposed a set of part-of-speech tags that handle most of token level phenomena of tweet. 
We are inspired by their work but argue that, first different abbreviation can have different syntactic functions. 
Like “mfw” is usually followed by an adverbial clause and “ima” is usually followed by a clausal complement. 
It is not reasonable to treat them in the same part-of-speech. 
In this paper, when annotating the POS tagging for abbreviations, we first try to recover their original forms, then use the POS of the core-word as the POS for the abbreviation.
Second, four special POS tags (S, L, M, Y) were designed to handle contraction words in Gimpel et al. \shortcite{Gimpel:2011:PTT:2002736.2002747}. Major concern of designing such tags is to minimize the effort of tokenization. 
However, contractions of common nouns and pronouns are casted into the same category which increase the difficulty of distinguishing their syntactic function (say, there's and book'll are treated with the same syntactic function). What's more, only a small proportion of words can be categorized into these tags (2.7 \% in total), which cast a doubt of the usefulness of these certain tags. In this paper, we believe such contraction can be properly handled by tokenization module, so we suggest to tokenize the contraction word and annotate POS tag accordingly.
Besides the contraction that be conventionally tokenized, tweets also witness a set of unconventional contraction like iv (I've), whatis (what is). In this paper, we follow the same idea of annotation abbreviation to handle the unconventional contractions and use the POS of core word of the original form as their POS.
Third, special POS was designed to handle emoticon in Gimpel et al. \shortcite{Gimpel:2011:PTT:2002736.2002747}. However, in most cases, emoticon plays the same role as most of the symbolic tokens. In this paper, we follow the UD guideline to annotate the emoticon as symbol (SYM).
At last, it’s arguable that some of the hashtags, URLs can work as a nominal in tweets. Whether treating them as the same part-of-speech or different ones according to their context is an open question. A preliminary survey on the standard UD English data shows that URL, email address are all tagged as the foreign language (X), so we also tag them as X and leave the disambiguation of their syntactic function to the annotation of parse tree.

We use the Universal POS tags~\cite{PETROV12.274} to tag the tweet tokens.

\subsection{Dependency Annotation}

@@

\subsection{POS Tagging}
